---
title: "Homework 1"
author: Safiya Sirota
date: 2021-09-29
output: 
  github_document:
    toc: TRUE
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
```

## Problem 1

__Creating a data frame:__

```{r}
set.seed(11)

problem1_df <-
  tibble(
    vec_numeric = rnorm(10),
    vec_logical = vec_numeric > 0,
    vec_char = character(10),
    vec_factor = factor(x = character(10), levels = 3) 
  )
```

__Taking the mean of variables in `problem1_df`:__

```{r, warning = FALSE}
mean(pull(problem1_df, vec_numeric))
mean(pull(problem1_df, vec_logical))
mean(pull(problem1_df, vec_char))
mean(pull(problem1_df, vec_factor))
```

It works to take the mean of the numeric and logical vectors. It doesn't work to take the mean of the character or factor variables. In fact, the warning given when trying to take the mean of a character or factor variable states that the argument is not numeric or logical, meaning the variables _must_ be numeric or logical to calculate a mean.

__Trying to convert variables to numeric type using `as.numeric`:__

```{r, eval = FALSE}
as.numeric(pull(problem1_df, vec_logical))
as.numeric(pull(problem1_df, vec_char))
as.numeric(pull(problem1_df, vec_factor))
```

When we use the `as.numeric` function on the logical variable, `vec_logical` gets converted to a binary variable where every value is either a 0 (representing FALSE) or a 1 (representing TRUE). When we use the `as.numeric` function on the character or factor variables, we get NA. 

When we take the mean of a logical variable, each T/F value is automatically converted to 0 or 1, and we take the mean of this numeric vector. We cannot compute the means of character or factor variables because they have no numeric conversion.

## Problem 2

__Loading penguins dataset:__

```{r}
data("penguins", package = "palmerpenguins")
```

__Describing `penguins`:__

* The variables in this dataset include:

  * __`r names(penguins)[1]`__ (Adelie, Chinstrap, or Gentoo),
  
  * __`r names(penguins)[2]`__ (Bisoe, Dream, or Torgersen),
  
  * __some body measurements__ (i.e., `r names(penguins)[3]`, `r names(penguins)[4]`, `r names(penguins)[5]`, and `r names(penguins)[6]`), 
  
  * __`r names(penguins)[7]`__, 
  
  * and __`r names(penguins)[8]`__ (2007, 2008, or 2009).

* The dataset has `r nrow(penguins)` observations and `r ncol(penguins)` variables, which are all listed above. 

* The mean flipper length of penguins in this dataset is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)` mm.

__Making a scatterplot:__

```{r, warning = FALSE, message = FALSE}
ggplot(penguins, aes(x = bill_length_mm , y = flipper_length_mm, color = species)) + geom_point()
```

```{r, eval = FALSE, echo = FALSE}
ggsave("hw1_scatterplot.png")
```

